Volunteers in the Clouds: A Web Based Architecture to Harness Volunteer Resources for Low Cost Distributed Evolutionary Computation

### 2015-12-12 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
NodIO-W in abstract
A sentence to talk about the Web Workers version.
Fixes Latex error
Closes #68, reference 63 is now 64 and new reference added.

\begin{abstract}
Volunteer computing is extensively used 
in science and industry, but it
still has a long way to go in order to become
 a mainstream
technology in the same way the cloud technologies in general have. The
lack of 
general purpose and open source tools with built-in analytics 
could be one of the reasons of why this is so, as well as the lack of
proofs of concept. That is why in this paper we present {\sf NodIO}, a
framework used to create client-server architectures for distributed
evolutionary algorithms, entirely written  
in JavaScript, which needs
minimal server support to work and is able to support persistent,
asynchronous, distributed evolutionary algorithms running in the
browser, as well as any other client consuming its REST application
programming interface.  The architecture, along with some initial
results on the expected performance and scaling behavior when the
number of users is increased is presented. We also use
the algorithm data to find out what kind of patterns the user behavior
follows by examining the resources they devote to our experiment, and
try to fit them to some statistical distribution to create the bases
of a model of user behavior in volunteer computing experiments of this
kind that can be used by researchers to predict its performance. 
In general, our experiments prove, first, that {\sf NodIO} is a flexible and
low-overhead platform to perform distributed computation experiments
that include volunteers, and then that user behavior follows some patterns,
and, thus, are amenable to a model.
\end{abstract}

@@

\subsubsection{Setup of Rastrigin's experiment in {\sf NodIO-W$^2$ }
This experiment was deployed in a virtual private server with 512MB of memory
running Ubuntu OS version 14.04 and hosted in DigitalOcean.com. Unlike previous
experiments where the optimum was expected to be found many times, 
this problem was expected to run constantly for many days finding only
sub-optimal approximations. In order to achieve this, the only change needed 
in the system was to limit the number of chromosomes kept in the pool. 
Previously, every PUT request which included a new chromosome increased 
the size of the pool, in a long-running experiment this has to be avoided
in order to not exhaust the available memory. In this experiment the size
of a chromosome was 1000 double precision numbers, considering this
and the memory available in the server, the the pool was limited 
to 10,000 chromosomes.  In order to keep this size
constant a random chromosome was pulled out of the pool before inserting a
new one, when the pool was full. For this problem {\sf NodEO}'s {\tt chromosome-float} 
and {\tt classic-ea-float} modules were used. The parameters for the EA 
were tournament size = 3 and  population size = 500, returning exchanging a
chromosome with the server every 100 generations. Again two workers in each window.
As before, a few participation calls were published in social networks.

>>>>

\begin{abstract}
Volunteer computing is extensively used 
in science and industry, but it
still has a long way to go in order to become
 a mainstream
technology in the same way the cloud technologies in general have. The
lack of 
general purpose and open source tools with built-in analytics 
could be one of the reasons of why this is so, as well as the lack of
proofs of concept. That is why in this paper we present {\sf NodIO}, a
framework used to create client-server architectures for distributed
evolutionary algorithms, entirely written  
in JavaScript, which needs
minimal server support to work and is able to support persistent,
asynchronous, distributed evolutionary algorithms running in the
browser, as well as any other client consuming its REST application
programming interface. An enhanced implementation, {\sf NodIO-W$^2$} is also
presented, this version leverages the HTML 5 Web Workers specification
to execute several clients per browser in the background, so that the the number of
generations per user could be improved. The architecture, along with some initial
results on the expected performance and scaling behavior when the
number of users is increased is presented. We also use
the algorithm data to find out what kind of patterns the user behavior
follows by examining the resources they devote to our experiment, and
try to fit them to some statistical distribution to create the bases
of a model of user behavior in volunteer computing experiments of this
kind that can be used by researchers to predict its performance. 
In general, our experiments prove, first, that {\sf NodIO} is a flexible and
low-overhead platform to perform distributed computation experiments
that include volunteers, and then that user behavior follows some patterns,
and, thus, are amenable to a model.
\end{abstract}

@@

\subsubsection{Setup of Rastrigin's experiment in {\sf NodIO-W$^2$ }}

This experiment was deployed in a virtual private server with 512MB of memory
running Ubuntu OS version 14.04 and hosted in DigitalOcean.com. Unlike previous
experiments where the optimum was expected to be found many times, 
this problem was expected to run constantly for many days finding only
sub-optimal approximations. In order to achieve this, the only change needed 
in the system was to limit the number of chromosomes kept in the pool. 
Previously, every PUT request which included a new chromosome increased 
the size of the pool, in a long-running experiment this has to be avoided
in order to not exhaust the available memory. In this experiment the size
of a chromosome was 1000 double precision numbers, considering this
and the memory available in the server, the the pool was limited 
to 10,000 chromosomes.  In order to keep this size
constant a random chromosome was pulled out of the pool before inserting a
new one, when the pool was full. For this problem {\sf NodEO}'s {\tt chromosome-float} 
and {\tt classic-ea-float} modules were used. The parameters for the EA 
were tournament size = 3 and  population size = 500, returning exchanging a
chromosome with the server every 100 generations. Again two workers in each window.
As before, a few participation calls were published in social networks.
